An equivalence of compositional frameworks

### 2019-05-30 EquivCompFrameworks_20180911kc.tex
rearranged and slightly polished appendix

\section{A symmetric monoidal double category of decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \mathbb{C}$sp mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} and then finally decategorify the obtained symmetric monoidal bicategory to obtain a symmetric monoidal category $F$Csp which is a generalization of Fong's original decorated cospan category \cite{Fong}. 
The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}.

@@

%\begin{comment}
\subsection{2-categories}\label{sec:2cats}
For standard 2-categorical material, we refer the reader to there and there.

Recall that a pseudofunctor $F\colon\bicat{A}\to\bicat{B}$ between bicategories $\bicat{A}$ and $\bicat{B}$ is functorial up to coherent natural isomorphism, namely for composable arrows we have $F(g\circ f)\cong Fg\circ Ff$ and $F(1_a)\cong1_{Fa}$ satisfying standard axioms. 
Given pseudofunctors $F,G\colon \bicat{A} \to \bicat{B}$, a \define{pseudonatural transformation} $\sigma$ consists of {\chris notation below needs fixing}
\begin{itemize}
\item for each object $a \in \bicat{A}$, a morphism $\sigma_a \colon F(a) \to G(a)$ in $\bicat{B}$ 
\item for every pair of objects $a$ and $b$ of $\mathbf{A}$, we have natural isomorphisms
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\mathbf{A}(a,b)$};
\node (B) at (2,0) {$\mathbf{X}(F(a),F(b))$};
\node (C) at (0,-1) {$\mathbf{X}(G(a),G(b))$};
\node (D) at (2,-1) {$\mathbf{X}(F(a),G(b))$};
\node (E) at (1,-0.5) {$\sigma_{a,b} \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(B) edge node[right]{$(\sigma_{b})_*$} (D)
(A) edge node[left]{$G$} (C)
(C) edge node[above]{$(\sigma_a)^*$} (D);
\end{tikzpicture}
\]
where $(\sigma_a)^*$ and $(\sigma_b)_*$ are the functors induced by precomposition and postcompositon, respectively. Thus for each morphism $f \colon a \to b$ in $\mathbf{A}$, we have an invertible 2-morphism $\sigma_f \colon G(f) \sigma_a \xrightarrow{\sim} \sigma_{b} F(f)$ in $\mathbf{X}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(a)$};
\node (B) at (1,0) {$F(b)$};
\node (C) at (0,-1) {$G(a)$};
\node (D) at (1,-1) {$G(b)$};
\node (E) at (0.5,-0.5) {$\sigma_f \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F(f)$} (B)
(B) edge node[right]{$\sigma_{b}$} (D)
(A) edge node[left]{$\sigma_a$} (C)
(C) edge node[above]{$G(f)$} (D);
\end{tikzpicture}
\]
which is compatible with composition and identities.
\end{itemize}
%such that for each composable pair of morphisms $f \colon a \to b$ and $g \colon b \to c$ of $\mathbf{A}$, the following diagrams commute:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$(G(g)G(f))\sigma_a$};
%\node (A') at (2,0.5) {$G(g)(G(f) \sigma_a)$};
%\node (B) at (0,-0.5) {$G(gf) \sigma_a$};
%\node (C) at (4,0.5) {$G(g) (\sigma_b F(f))$};
%\node (C') at (4,-0.5) {$\sigma_c F(gf)$};
%\node (D) at (6,0.5) {$(G(g) \sigma_b) F(f)$};
%\node (D') at (8,-0.5) {$\sigma_c (F(g)F(f))$};
%\node (F) at (8,0.5) {$(\sigma_c F(g)) F(f)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$a^\prime$} (A')
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(A') edge node[above]{$1_{G(g)} \sigma_f$} (C)
%(B) edge node[above]{$\sigma_{gf}$} (C')
%(C) edge node [above] {${a^\prime}^{-1}$} (D)
%(D') edge node [above] {$1_{\sigma_c} \phi$} (C')
%(D) edge node [above] {$\sigma_g 1_{F(f)}$} (F)
%(F) edge node [right] {$a^\prime$} (D');
%\end{tikzpicture}
%\]
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$1_{G(a)} \sigma_a$};
%\node (B) at (1,1) {$G(1_a) \sigma_a$};
%\node (C) at (2,0) {$\sigma_a F(1_a)$};
%\node (D) at (0.5,-1) {$\sigma_a$};
%\node (E) at (1.5,-1) {$\sigma_a 1_{F(a)}$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(B) edge node[right]{$\sigma_{1_a}$} (C)
%(A) edge node [left] {$\ell^\prime$} (D)
%(D) edge node [above] {${\rho^\prime}^{-1}$} (E)
%(E) edge node [right] {$1_{\sigma_a} \phi$} (C);
%\end{tikzpicture}
%\]
%Christina: I removed the above, they are not needed I believe
We denote by $[\A,\bicat{Cat}]_\pse$ the 2-category 
of pseudofunctors, pseudonatural transformations and modifications from an ordinary category $\A$ viewed as a 2-category with trivial 2-cells, into $\bicat{Cat}$. Classicaly this has been referred to as the 2-category of \emph{opindexed categories}, since an indexed category is a contravariant pseudofunctor into $\bicat{Cat}$.

A \define{lax monoidal} pseudofunctor is....
A \define{symmetric lax monoidal} pseudofunctor...
{\chris Not sure if we need monoidal transformations?}

\subsection{Double categories}\label{sec:doublecats}
{\chris Moved from earlier, probably merge with what follows}
Once again following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
\[
  \xymatrix@-.5pc{
    A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
    B\ar[d]^g\\
    C \ar[r]|{|}_N & D
  }
\]
We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \colon \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \colon _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \colon \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \colon x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \colon a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}
\begin{defn}
A double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{strong} if the comparison and unit constraints are globular isomorphisms, meaning that for each composable pair of horizontal 1-cells $M$ and $N$ we have a natural isomorphism $$\mathbb{F}_{M,N} \colon \mathbb{F}(M) \odot \mathbb{F}(N) \xrightarrow{\sim} \mathbb{F}(M \odot N)$$and for each object $a \in \mathbb{A}$ a natural isomorphism $$\mathbb{F}_a \colon \hat{U}_{\mathbb{F}(a)} \xrightarrow{\sim} \mathbb{F}(U_a).$$
\end{defn}

{\chris Here ends what I moved}

@@

@@

\bibitem{CourPhD} K.\ Courser, Cospans for Networks, PhD Thesis, University of California at Riverside (2019)

%\bibitem{Ehresmann63} C.\ Ehresmann, Cat\'egories structur\'ees III: Quintettes et applications covariantes,  \textsl{Cah.\ Top.\ G\'eom.\ Diff.} \textbf{5} (1963), 1--22.

@@

>>>>

\section{A symmetric monoidal double category of decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \mathbb{C}$sp mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} and then finally decategorify the obtained symmetric monoidal bicategory to obtain a symmetric monoidal category $F$Csp which is a generalization of Fong's original decorated cospan category \cite{Fong}. 
The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of (pseudo) double categories are recalled in \cref{sec:doublecats}.

@@

%\begin{comment}

{\chris Here, I basically rearranged things to make them come in a certain order, and also added a few things for monoidal bicats, maybe more is needed or not. Didn't change Kenny's notation or writings almost at all}

In this chapter, we gather some well-known concepts required to make the material self-contained, as well as references to more detailed expositions.

\subsection{2-categories}\label{sec:2cats}
For standard 2-categorical material, we refer the reader to e.g. \cite{KS} or \cite{DS} for monoidal structures.

Recall that a pseudofunctor $F\colon\bicat{A}\to\bicat{B}$ between bicategories $\bicat{A}$ and $\bicat{B}$ is functorial up to coherent natural isomorphism, namely for composable arrows we have $F(g\circ f)\cong Fg\circ Ff$ and $F(1_a)\cong1_{Fa}$ satisfying standard axioms. 
Given pseudofunctors $F,G\colon \bicat{A} \to \bicat{B}$, a \define{pseudonatural transformation} $\sigma$ consists of
\begin{itemize}
\item for each object $a \in \bicat{A}$, a morphism $\sigma_a \colon F(a) \to G(a)$ in $\bicat{B}$ 
\item 
%for every pair of objects $a$ and $b$ of $\mathbf{A}$, we have natural isomorphisms
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$\mathbf{A}(a,b)$};
%\node (B) at (2,0) {$\mathbf{X}(F(a),F(b))$};
%\node (C) at (0,-1) {$\mathbf{X}(G(a),G(b))$};
%\node (D) at (2,-1) {$\mathbf{X}(F(a),G(b))$};
%\node (E) at (1,-0.5) {$\sigma_{a,b} \Nearrow$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$F$} (B)
%(B) edge node[right]{$(\sigma_{b})_*$} (D)
%(A) edge node[left]{$G$} (C)
%(C) edge node[above]{$(\sigma_a)^*$} (D);
%\end{tikzpicture}
%\]
%where $(\sigma_a)^*$ and $(\sigma_b)_*$ are the functors induced by precomposition and postcompositon, respectively. Thus 
for each morphism $f \colon a \to b$ in $\mathbf{A}$, an invertible natural 2-morphism $\sigma_f \colon G(f) \sigma_a \xrightarrow{\sim} \sigma_{b} F(f)$ in $\mathbf{B}$
%:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$F(a)$};
%\node (B) at (1,0) {$F(b)$};
%\node (C) at (0,-1) {$G(a)$};
%\node (D) at (1,-1) {$G(b)$};
%\node (E) at (0.5,-0.5) {$\sigma_f \Nearrow$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$F(f)$} (B)
%(B) edge node[right]{$\sigma_{b}$} (D)
%(A) edge node[left]{$\sigma_a$} (C)
%(C) edge node[above]{$G(f)$} (D);
%\end{tikzpicture}
%\]
compatible with composition and identities.
\end{itemize}
%such that for each composable pair of morphisms $f \colon a \to b$ and $g \colon b \to c$ of $\mathbf{A}$, the following diagrams commute:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$(G(g)G(f))\sigma_a$};
%\node (A') at (2,0.5) {$G(g)(G(f) \sigma_a)$};
%\node (B) at (0,-0.5) {$G(gf) \sigma_a$};
%\node (C) at (4,0.5) {$G(g) (\sigma_b F(f))$};
%\node (C') at (4,-0.5) {$\sigma_c F(gf)$};
%\node (D) at (6,0.5) {$(G(g) \sigma_b) F(f)$};
%\node (D') at (8,-0.5) {$\sigma_c (F(g)F(f))$};
%\node (F) at (8,0.5) {$(\sigma_c F(g)) F(f)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$a^\prime$} (A')
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(A') edge node[above]{$1_{G(g)} \sigma_f$} (C)
%(B) edge node[above]{$\sigma_{gf}$} (C')
%(C) edge node [above] {${a^\prime}^{-1}$} (D)
%(D') edge node [above] {$1_{\sigma_c} \phi$} (C')
%(D) edge node [above] {$\sigma_g 1_{F(f)}$} (F)
%(F) edge node [right] {$a^\prime$} (D');
%\end{tikzpicture}
%\]
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$1_{G(a)} \sigma_a$};
%\node (B) at (1,1) {$G(1_a) \sigma_a$};
%\node (C) at (2,0) {$\sigma_a F(1_a)$};
%\node (D) at (0.5,-1) {$\sigma_a$};
%\node (E) at (1.5,-1) {$\sigma_a 1_{F(a)}$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(B) edge node[right]{$\sigma_{1_a}$} (C)
%(A) edge node [left] {$\ell^\prime$} (D)
%(D) edge node [above] {${\rho^\prime}^{-1}$} (E)
%(E) edge node [right] {$1_{\sigma_a} \phi$} (C);
%\end{tikzpicture}
%\]
%Christina: I removed the above, they are not needed I believe
We denote by $[\A,\bicat{Cat}]_\pse$ the 2-category 
of pseudofunctors, pseudonatural transformations and modifications from an ordinary category $\A$ viewed as a 2-category with trivial 2-cells, into $\bicat{Cat}$. This is also referred to as the 2-category of \emph{opindexed categories}, since an indexed category is a contravariant pseudofunctor into $\bicat{Cat}$.

A \define{lax monoidal} pseudofunctor between monoidal bicategories $F\colon\bicat{A}\to\bicat{B}$ is a pseudofunctor equipped with pseudonatural transformations with components $\mu_{a,b}\colon Fa\otimes Fb\to F(a\otimes b)$ and $\mu_0\colon I\to FI$ along with coherent invertible modifications for associativity and unitality. This is commonly also called `weak monoida' pseudofunctor. A \define{symmetric lax monoidal} pseudofunctor between symmetric monoidal bicategories comes with invertible $F(b)\circ \mu_{a,b}\cong \mu_{b,a}\circ b$, where $b$ is the braiding. 
{\chris Hiding various details under the rug - e.g. what a symmetric monoidal bicategory is. Should be fine? Also, not sure if we need monoidal transformations?}

\subsection{Double categories}\label{sec:doublecats}

@@

{\chris Pretty sure the def below can be incorporated above, however got a bit lost in different notation. Kenny Fix?}
\begin{defn}
A double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{strong} if the comparison and unit constraints are globular isomorphisms, meaning that for each composable pair of horizontal 1-cells $M$ and $N$ we have a natural isomorphism $$\mathbb{F}_{M,N} \colon \mathbb{F}(M) \odot \mathbb{F}(N) \xrightarrow{\sim} \mathbb{F}(M \odot N)$$and for each object $a \in \mathbb{A}$ a natural isomorphism $$\mathbb{F}_a \colon \hat{U}_{\mathbb{F}(a)} \xrightarrow{\sim} \mathbb{F}(U_a).$$
\end{defn}

Following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:
\[
  \xymatrix@-.5pc{
    A \ar[r]|{|}^{M}  \ar[d]_f \ar@{}[dr]|{\Downarrow a}&
    B\ar[d]^g\\
    C \ar[r]|{|}_N & D
  }
\]
We call $M$ and $N$ the \define{horizontal source and target} of the 2-morphism $a$, respectively, and likewise we call $f$ and $g$ the \define{vertical source and target} of the 2-morphism $a$, respectively. Thus $_f \mathbb{A}_g(M,N)$ denotes the set of 2-morphisms in $\mathbb{A}$ with horizontal source and target $M$ and $N$ and vertical source and target $f$ and $g$.
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{full} (respectively, \define{faithful}) if $\mathbb{F}_0 \colon \mathbb{A}_0 \to \mathbb{X}_0$ is full (respectively, faithful) and each map $$\mathbb{F}_1 \colon _f \mathbb{A}_g(M,N) \to _{\mathbb{F}(f)} \mathbb{X}_{\mathbb{F}(g)}(\mathbb{F}(M),\mathbb{F}(N))$$ is surjective (respectively, injective).
\end{defn}
\begin{defn}
A (possibly lax or oplax) double functor $\mathbb{F} \colon \mathbb{A} \to \mathbb{X}$ is \define{essentially surjective} if we can simultaneously make the following choices:
\begin{enumerate}
\item{For each object $x \in \mathbb{X}$, we can find an object $a \in \mathbb{A}$ together with a vertical 1-isomorphism $\alpha_x \colon \mathbb{F}(a) \to x$, and}
\item{For each horizontal 1-cell $N \colon x_1 \tobar x_2$  of $\mathbb{X}$, we can find a horizontal 1-cell $M \colon a_1 \tobar a_2$ of $\mathbb{A}$ and a 2-isomorphism $a_{N}$ of $\mathbb{X}$ as in the following diagram:
\[
  \xymatrix@-.5pc{
    \mathbb{F}(a_1) \ar[r]|{|}^{\mathbb{F}(M)}  \ar[d]_{\alpha_{x_1}} \ar@{}[dr]|{\Downarrow a_N}&
    \mathbb{F}(a_1) \ar[d]^{\alpha_{x_2}}\\
    x_1 \ar[r]|{|}_N & x_2
  }
\]
}
\end{enumerate}
\end{defn}

@@

\bibitem{CourPhD} K.\ Courser, Open Systems: a Double Categorical Perspective, PhD Thesis, University of California at Riverside (2019)

\bibitem{DS} B.\ Day and R.\ Street, Monoidal Bicategories and Hopf Algebroids, \emph{Adv. Math.}, 129(1):99--157, 1997.
%\bibitem{Ehresmann63} C.\ Ehresmann, Cat\'egories structur\'ees III: Quintettes et applications covariantes,  \textsl{Cah.\ Top.\ G\'eom.\ Diff.} \textbf{5} (1963), 1--22.

@@

\bibitem{KS} G.\ Kelly and R.\ Street, Review of the elements of 2-categories, \emph{Lecture Notes in Mathematics} vol. 40, pages 75---103, Springer Berlin/Heidelberg, 1974.
